%% ----------------------------------------------------------------
%% Thesis.tex -- MAIN FILE (the one that you compile with LaTeX)
%% ----------------------------------------------------------------
% Set up the document
\documentclass[a4paper, 11pt, oneside]{Thesis}  % Use the "Thesis" style, based on the ECS Thesis style by Steve Gunn
\graphicspath{Figures/}  % Location of the graphics files (set up for graphics to be in PDF format)
% Include any extra LaTeX packages required
\usepackage{changepage}
\usepackage{fancyhdr}
\usepackage{float}
\usepackage{import}
\usepackage{lipsum}
\usepackage[square, numbers, comma, sort&compress]{natbib}  % Use the "Natbib" style for the references in the Bibliography
\usepackage{setspace}
\hypersetup{urlcolor=blue, colorlinks=true}  % Colours hyperlinks in blue, but this can be distracting if there are many links.
\usepackage{vector}  % Allows "\bvec{}" and "\buvec{}" for "blackboard" style bold vectors in maths
\usepackage{verbatim}  % Needed for the "comment" environment to make LaTeX comments
\usepackage{subfig}
\usepackage{graphicx}
\usepackage{gensymb}
\usepackage{afterpage, geometry, pdflscape, tabularx, ulem} % for landscape tables
\usepackage{multirow}
\usepackage{pdfpages}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{afterpage}

\definecolor{ForestGreen}{RGB}{34, 139, 34}
\definecolor{Rhodamine}{RGB}{188, 82, 134}

\newgeometry{left=2in,right=0in,bottom=0in,top=2in}

% Use this to change the margin width for a snippet of text/table/figure etc.
\def\changemargin#1#2{\list{}{\rightmargin#2\leftmargin#1}\item[]}
\let\endchangemargin=\endlist

%\def \thesistitle{Sensing and Actuation in Conductive Particle Electroactive Elastomeric Bodies}
\def \thesistitle{Integration of Biomimetic Pressure Mapping and Actuation in Soft Electroactive Elastomer Composites}
%\def \thesistitle{Integrated sensing and actuation in stretchable devices - EIT and EAP integration}
%\def \thesistitle{Electrical impedance tomography based pressure mapping integration with dielectric elastomer actuators using conductive particle elastomer composites.}
%\def \thesistitle{Synergy Between Sensing and Actuation: Electroactive Elastomeric Composite Bodies for Soft Robotics.}
%\def \thesistitle{The Interplay of Two Soft Robotics Sensor and Actuator Technologies using Carbon Based Elastomer Composites.}
%\def \thesistitle{Beyond Rigid: Forming and Characterising Novel Soft Sensor-Actuator Technologies.}
%\def \thesistitle{Towards Artificial Skin and Muscle: Soft Electroactive Elastomeric Bodies that Can Sense and Contract}

%% ----------------------------------------------------------------
\begin{document}
	\frontmatter      % Begin Roman style (i, ii, iii, iv...) page numbering
	% Set up the Title Page
	\title  {\thesistitle}
	\authors  {\texorpdfstring
		{\href{ellingham.richie@gmail.com}{Richard James Morrin Ellingham}}
		{Richard James Morrin Ellingham}
	}
	\addresses  {\groupname\\\deptname\\\univname}  % Do not change this here, instead these must be set in the "Thesis.cls" file, please look through it instead
	\date       {\today}
	\subject    {}
	\keywords   {}
	\maketitle
	%% ----------------------------------------------------------------
	\setstretch{1.5}  % It is better to have smaller font and larger line spacing than the other way round
	% Define the page headers using the FancyHdr package and set up for one-sided printing
	\fancyhead{}  % Clears all page headers and footers
	\rhead{\thepage}  % Sets the right side header to show the page number
	\lhead{}  % Clears the left side page header
	\pagestyle{fancy}  % Finally, use the "fancy" page style to implement the FancyHdr headers

	% The "Funny Quote Page"
	\pagestyle{empty}  % No headers or footers for the following pages
	\null\vfill
	% Now comes the "Funny Quote", written in italics
%	\textit{``PHD stands for Permanent Head Damage''}
	\textit{``We've made good progress with the brains of robots but their bodies are still primitive. We need a new kind of robot body that is inspired by elegance, efficiency, and by the soft materials found in nature. Indeed, this has become the key idea of a new field of research called soft robotics.''}
	\begin{flushright}
		C. Keplinger, 2019
	\end{flushright}
	\vfill\vfill\vfill\vfill\vfill\vfill\null
	\clearpage  % Funny Quote page ended, start a new page
	\setstretch{1.0}
	%% ----------------------------------------------------------------
	% The Abstract Page
	\addtotoc{Abstract}  % Add the "Abstract" page entry to the Contents
	\abstract{\addtocontents{toc}{\vspace{1em}}  % Add a gap in the Contents, for aesthetics
		% 
	Some of the world's most advanced technology is rigid due to various factors such as; manufacturability, miniaturisability, physical linearity, and more ideal linear physical mechanics. Since before the industrial revolution rigid sensors and actuators have been evolving and have been refined to a high level, but now these rigid devices are encountering limitations, leading to demand for soft robotic solutions that mimic biological evolution. This thesis develops and integrates an artificial skin and artificial muscle technology by combining of Electrical Impedance Tomography (EIT), carbon black silicone rubber (CBSR) elastomer composites, and Dielectric Elastomer Actuators (DEAs) to enhance pressure mapping, strain sensing, and actuation capabilities. 
	CBSR elastomer composites, noted for their high stretchability, piezoresistivity, and biocompatibility, were investigated to understand their non-linear time dependent phenomena in particular, resistance relaxation behavior. This research contributes to optimising the design of flexible dynamic strain sensors by modelling the response of resistance to transient strain inputs. The next part of the thesis describes development of an EIT-based pressure mapping system using a CBSR sensing domain that mimics pressure mapping qualities of human skin. This system was evaluated for its spatial and temporal resolution, showing potential for creating artificial pressure-sensitive skin with practical applications. At the core of this thesis, the integration of EIT-based sensing with DEAs was tested and analysed to improve the mapping of compressive forces across the actuator electrode surfaces. Despite trade-offs in accuracy due to electrode compliance, this approach offers promising advancements for applications requiring precise actuation and pressure mapping. This work has contributed towards filing a patent for an DEA-EIT actuator-sensor device. Additionally, the research has uncovered unintentional power generation in DEA-EIT systems through FEA studies and analytical validation, which could function as Dielectric Elastomer Generators (DEGs) with mapped compressive load inputs. This finding highlights the multi-functionality of DEAs and suggests opportunities for energy harvesting applications. To aid the portability of EIT-based pressure mapping sensor developed, a low-cost circuit and evaluation tools were developed. This system enables comprehensive characterisation of various sensing domains and supports advancements in EIT-based soft sensor technology, with implications for biomedical devices, robotics, and energy harvesting. 
	Overall, this thesis advances the field of soft sensors by investigating an electroactive material and using the material to integrate EIT-based pressure mapping with dielectric elastomer actuator technologies, providing new insights and applications in dynamic sensing and actuation.
	}
	\clearpage  % Abstract ended, start a new page
	%% ----------------------------------------------------------------

	%% ----------------------------------------------------------------
	% Declaration Page required for the Thesis, your institution may give you a different text to place here
	\Declaration{
		\addtocontents{toc}{\vspace{1em}}  % Add a gap in the Contents, for aesthetics
		This form is to accompany the submission of any thesis that contains research reported in co-authored work that has been published, accepted for publication, or submitted for publication. A copy of this form should be included for each co-authored work that is included in the thesis.	Completed forms should be included at the front (after the thesis abstract) of each copy of the thesis submitted for examination and library deposit.
		
		\textbf{This form should be completed by a co-author, not by the candidate.}
		
		Please indicate the chapter/section/pages of this thesis that are extracted from co-authored work and provide	details of the publication or submission from the extract comes:
		\begin{enumerate}
			\item ``Stress and Resistance Relaxation for Carbon Nanoparticle Silicone Rubber Composite Large-Strain Sensors'' - Proceedings 17th IEEE/ASME International Conference on Mechatronic and Embedded Systems and Applications (MESA) - Ellingham R, Giffney T, 2021. 
			
			Data and analysis from this work is present throughout \textbf{Chapter 3}.
			
			\item ``Carbon Black Silicone Piezoresistive Electrical Impedance Tomography Stress Sensor Device'' -  Proceedings Electroactive Polymer Actuators and Devices (EAPAD) XXIV - Ellingham R, Giffney T, 2022.
			
			The methodology from this work is present \textbf{Chapter 4}. 
			
			\item ``An Electrical Impedance Tomography Based Artificial Soft Skin Pressure Sensor: Characterisation and Force Modelling'' - Sensors and Actuators A: Physical - Ellingham R, Pretty C, Holder-Pearson L, Aw K, Giffney T, 2024.

			\textbf{Chapter 4} is an adaptation of this work.
			
			\item ``Deformation Mapping in Dielectric Elastomer Actuators using Electrical Impedance Tomography'' - Proceedings Electroactive Polymer Actuators and Devices (EAPAD) XXVI - Ellingham R, Choi Y, Giffney T, 2024.
			
			\textbf{Chapter 6} is an extension of this work.
			
			\item ``A Portable Electrical Impedance Tomography Based Pressure Mapping Sensor and Validation System'' - HardwareX - Ellingham R, Holder-Pearson L, Pretty C, Giffney T, [Under Review, Submitted Aug 2024]. 
			
			\textbf{Chapter 5} is adpated from this work.
			
		\end{enumerate}
		
		Please detail the nature and extent (\%) of contribution by the candidate:
		\begin{enumerate}
			\item The methodology, hardware, and software were all developed by the candidate. Experimentation and analysis of results and written content were produced by the candidate. All other authors reviewed the draft manuscript. 90\% of the research and 90\% of the writing were contributed by the candidate.
			\item The methodology, hardware, and software were all developed by the candidate. Experimentation and analysis of results and written content were produced by the candidate. All other authors reviewed the draft manuscript. 90\% of the research and 90\% of the writing were contributed by the candidate.
			\item The methodology, hardware, and software were all developed by the candidate. Experimentation and analysis of results and written content were produced by the candidate. All other authors reviewed the draft manuscript. 90\% of the research and 90\% of the writing were contributed by the candidate.
			\item The methodology, hardware, and software were all developed by the candidate. Experimentation and analysis of results and written content were produced by the candidate. All other authors reviewed the draft manuscript. 90\% of the research and 90\% of the writing were contributed by the candidate.
			\item The methodology, hardware, and software were all developed by the candidate. Experimentation and analysis of results and written content were produced by the candidate. All other authors reviewed the draft manuscript. 90\% of the research and 90\% of the writing were contributed by the candidate.
		\end{enumerate}
	
		\textbf{Certification by Co-authors:}
		
		If there is more than one co-author then a single co-author can sign on behalf of all.
		
		The undersigned certifies that:
		\begin{itemize}
			\item The above statement correctly reflects the nature and extent of the Doctoral candidate’s contribution to this co-authored work
			\item In cases where the candidate was the lead author of the co-authored work he or she wrote the text
		\end{itemize}
			
		Name: Tim Giffney \\
		\hspace{1cm}\rule[1em]{20em}{0.5pt} \\  % This prints a line for the signature
		Signed: \textit{Tim Giffney}\\
		\hspace{1cm}\rule[1em]{20em}{0.5pt} \\  % This prints a line for the signature
		Date: \textit{30-10-2024}\\
		\hspace{1cm}\rule[1em]{20em}{0.5pt} \\ % This prints a line to write the date
	}
	\clearpage  % Declaration ended, now start a new page
	
	\setstretch{1.3}  % Reset the line-spacing to 1.3 for body text (if it has changed)
	%% ----------------------------------------------------------------
	% The Acknowledgements page, for thanking everyone
	\acknowledgements{
		\addtocontents{toc}{\vspace{1em}}  % Add a gap in the Contents, for aesthetics
	I would like to thank many people who have implicitly and explicitly helped me get through my doctorate. First, the UC Bioengineering office including the Lung Ladies, Lung lads, IPC, Hamish for infinite Friday beers, and Nic's abstract art. Also included in the office, Geoff Chase, for a constant stream of stories and advice always within earshot. The Mechanical Engineering post-grad office for baking, fruit, and comic relief. The UC Mechanical and Electrical department technicians for yarns and practical advice. All of the students and project groups I tutored and mentored, keeping me on my toes throughout the doctorate. Thanks for the motivation from all of my friends with the never-ending question `so... when do you finish', a very motivating and difficult question answer. Something that prolonged the journey but kept me sane, were the technical the side jobs and fun adventures with colleagues and friends maintaining my sanity. A special thanks to my supervision and collaborations abroad, including my materials science work in IMDEA with Juan Jose and Anastasia and my time with with the MEiTNER soft robotics research group led by Markus Vorrath. Thanks to my internal supervision from Tim teaching me what PHD stands for and my co-supervisors Chris and Lui who have been a good source of advice, yarns, and general guidance. A thank you to my family for being supportive in times of strife. Finally, I would like to thank my partner Tess for her ongoing support and distractions from my PhD work.
	}
	\clearpage  % End of the Acknowledgements
	%% ----------------------------------------------------------------
	\pagestyle{fancy}  %The page style headers have been "empty" all this time, now use the "fancy" headers as defined before to bring them back
	%% ----------------------------------------------------------------
	\lhead{\emph{Contents}}  % Set the left side page header to "Contents"
	\tableofcontents  % Write out the Table of Contents
	%% ----------------------------------------------------------------
	\lhead{\emph{List of Figures}}  % Set the left side page header to "List if Figures"
	\listoffigures  % Write out the List of Figures
	%% ----------------------------------------------------------------
	\lhead{\emph{List of Tables}}  % Set the left side page header to "List of Tables"
	\listoftables  % Write out the List of Tables
	%% ----------------------------------------------------------------
	\setstretch{1.5}  % Set the line spacing to 1.5, this makes the following tables easier to read
	\clearpage  % Start a new page
	\vspace{-1cm}
	\lhead{\emph{Abbreviations}}  % Set the left side page header to "Abbreviations"
	\vspace{-0.5cm}
	\listofsymbols{ll}  % Include a list of Abbreviations (a table of two columns)
	{
		% \textbf{Acronym} & \textbf{W}hat (it) \textbf{S}tands \textbf{F}or \\
		\textbf{ADC} & \textbf{A}nalog-\textbf{t}o-\textbf{D}igital \textbf{C}onverter \\
		\textbf{CAD} & \textbf{C}omputer \textbf{A}ided \textbf{D}esign \\
		\textbf{CB} & \textbf{C}arbon \textbf{B}lack \\
		\textbf{CE} & \textbf{C}ompliant \textbf{E}electrode \\
		\textbf{CFA} & \textbf{C}artesian \textbf{F}orce \textbf{A}pplicator \\
		\textbf{CoM} & \textbf{C}enter of \textbf{M}ass \\
		\textbf{CPEC} & \textbf{C}onductive \textbf{P}article \textbf{E}lastomer \textbf{C}omposite\\
		\textbf{DE} & \textbf{D}ielectric \textbf{E}lastomer \\
		\textbf{DEA} & \textbf{D}ielectric \textbf{E}lastomer \textbf{A}ctuator \\
		\textbf{DEG} & \textbf{D}ielectric \textbf{E}lastomer \textbf{G}enerator \\
		\textbf{DUT} & \textbf{D}omain \textbf{U}nder \textbf{T}est \\
		\textbf{EIT} & \textbf{E}lectrical \textbf{I}mpedance \textbf{T}omography \\
		\textbf{ERT} & \textbf{E}lectrical \textbf{R}esistance \textbf{T}omography \\
		\textbf{FEA} & \textbf{F}inite \textbf{E}lement \textbf{A}nalysis \\
		\textbf{FEM} & \textbf{F}inite \textbf{E}lement \textbf{M}odelling \\
		\textbf{FPC} & \textbf{F}lexible \textbf{P}rinted \textbf{C}ircuit \\
		\textbf{MUX} & \textbf{M}ultiplexer \\
		\textbf{PCB} & \textbf{P}rinted \textbf{C}ircuit \textbf{B}oard \\
		\textbf{PCBA} & \textbf{P}rinted \textbf{C}ircuit \textbf{B}oard \textbf{A}ssembly\\
		\textbf{PDMS} & \textbf{P}oly\textbf{d}i\textbf{m}ethyl\textbf{s}iloxane (AKA silicone)\\
		\textbf{PNEC} & \textbf{P}iezoresistive \textbf{N}anoparticle \textbf{E}lastomer \textbf{C}omposite\\
		\textbf{SMU} & \textbf{S}ource \textbf{M}easure \textbf{U}nit \\
		\textbf{SMD} & \textbf{S}urface-\textbf{M}ount \textbf{D}evice \\
		\textbf{SR} & \textbf{S}ilicone \textbf{R}ubber \\
		\textbf{EAP} & \textbf{E}lectro-\textbf{A}ctive \textbf{P}olymer\\

	}
	%% ----------------------------------------------------------------
	
	\clearpage  %Start a new page
	\lhead{\emph{Nomenclature}}  % Set the left side page header to "Symbols"
	\listofnomenclature{lll}  % Include a list of Symbols (a three column table)
	{
		% symbol & name & unit \\
		$A$ & Area & [m$^2$] \\
		$C$ & Capacitance & [F] \\
		$\epsilon_0$ & Vacuum Permittivity & [Fm$^{-1}$] \\
		$\epsilon_r$ & Relative Permittivity & [a.u.] \\
		$K$ & Bulk Modulus & [Pa] \\
		$\nu$ & Poisson’s Ratio & [a.u.] \\
		$Q$ & Electrical Charge & [C] \\
		$U_E$ & Electrical Potential Energy & [J] \\
		$U_{\epsilon}$ & Elastic Potential Energy & [J] \\
		$R$ & Resistance & [$\Omega$] \\
		$\sigma$ & Stress & [Pa] \\
		$S$/$\varepsilon$ & Strain & [a.u. or \%] \\ % change this to S or \varepsilon?
		$\dot\varepsilon$ & Strain Rate & [\%s$^{-1}$] \\
		$V$ & Voltage & [V] \\
		$Y$/$E$ & Young's Modulus & [Pa] \\
		$z$/$d$ & Thickness & [m] \\

		
	}
	%% ----------------------------------------------------------------
	% End of the pre-able, contents and lists of things
	% Begin the Dedication page
	\setstretch{1.5}  % Return the line spacing back
%	\pagestyle{empty}  % Page style needs to be empty for this page
%	\dedicatory{Dedicated to tinned baked beans in all their glory\ldots}
%	\addtocontents{toc}{\vspace{2em}}  % Add a gap in the Contents, for aesthetics
	%% ----------------------------------------------------------------
	\mainmatter	  % Begin normal, numeric (1,2,3...) page numbering
	\pagestyle{fancy}  % Return the page headers back to the "fancy" style
	% Include the chapters of the thesis, as separate files
	% Just uncomment the lines as you write the chapters
	\lhead{\emph{\chapiname}}
	\import{Chapters/}{Chapter1.tex} % Introduction
	\lhead{\emph{\chapiiname}}
	\import{Chapters/}{Chapter2.tex} % Literature Review
	\lhead{\emph{\chapiiiname}}
	\import{Chapters/}{Chapter3.tex} % CBSR behaviour
	\lhead{\emph{\chapivname}}
	\import{Chapters/}{Chapter4.tex} % Improved EIT-based sensor
	\lhead{\emph{\chapvname}}
	\import{Chapters/}{Chapter5.tex} % Portable EIT hardware 
	\lhead{\emph{\chapviname}}
	\import{Chapters/}{Chapter6.tex} % DEA-EIT
	\lhead{\emph{\chapviiname}}
	\import{Chapters/}{Chapter7.tex} % DEA-EIT an unintended DEG
	\lhead{\emph{\chapviiiname}}
	\import{Chapters/}{Chapter8.tex} % Discussion and Conclusion

	%% ----------------------------------------------------------------
	% Now begin the Appendices, including them as separate files
	\addtocontents{toc}{\vspace{2em}} % Add a gap in the Contents, for aesthetics
	\appendix % Cue to tell LaTeX that the following 'chapters' are Appendices
%	\input{Appendices/AppendixA}
	\input{Appendices/AppendixB}
	\input{Appendices/AppendixC}
	\input{Appendices/AppendixD}
	\input{Appendices/AppendixE}
	\input{Appendices/AppendixF}
		
	\addtocontents{toc}{\vspace{2em}}  % Add a gap in the Contents, for aesthetics
	\backmatter
	%% ----------------------------------------------------------------
	\label{Bibliography}
	\lhead{\emph{Bibliography}}  % Change the left side page header to "Bibliography"
	\bibliographystyle{unsrtnat}  % Use the "unsrtnat" BibTeX style for formatting the Bibliography
	\bibliography{Bibliography}  % The references (bibliography) information are stored in the file named "Bibliography.bib"
\end{document}  % The End
%% ----------------------------------------------------------------