\chapter{\chapviiiname}
\label{chapter8}
%Here we are. The final chapter. Time to reflect and get philosophical!
A literature review comparing human skin and muscle tissue to current soft pressure mapping and soft electroactive actuator technologies was completed. From this review it was determined that CBSR material would be suitable for a range of soft pressure mapping and actuation devices because it is readily available, non-toxic, conductive, piezoresistive, mouldable, and highly deformable. Two specific technologies that emulate skin in terms of pressure mapping and muscle in terms of actuation were chosen and combined; EIT-based pressure mapping and dielectric elastomer actuators. Both sensor and actuator technologies use CBSR material as the electroactive polymer (EAP) component.

The first novel work in this thesis explored the CPEC material characterisation and selection of carbon black (CB) silicone rubber (SR) composites through several imaging techniques, including microscopy, scanning electron microscopy, and Raman spectroscopy. This imaging showed the carbon black micro-structure and how it was dispersed through a the micro-void filled elastomer composite. Then a macro-characterisation of the material was completed by determining several repeatable dynamic and transient characteristics by fitting several mathematical formulae to the repeated phenomena. This characterisation shows that the material has repeatable characteristics and elucidated how the material reacts to different time series tensile strain inputs, setting a foundation for development of sensors and actuators from the composite.

Using different configurations of the same CBSR material in Chapter \ref{chapter3}, electrical impedance tomography (EIT) was used to map dynamic changes in resistance within a CBSR sheet of material undergoing a series of compressive loads. A set of performance metrics were developed as tools to determine the quality of a reconstruction given a known load input. The resistive relaxation was quantified in the 2D EIT reconstruction and compared to a 1D equivalent resistive relaxation to show that 1D data and characterisation could be extrapolated to handle 2D problems. A model was then created which translates changes in conductivity to force for elements within the 2D image reconstruction. 

To allow EIT-based pressure mapping to be used in a variety of portable/wearable, small scale applications a open-source circuit was developed. In parallel, a Cartesian force applicator was developed especially to capture compressive load data so that such pressure sensor could be calibrated and characterised for a range of different sensor domain materials.

Dielectric elastomer actuators are soft actuators that can emulate human muscle tissue function. To enhance the functionality of these actuator devices dielectric elastomer actuator technology was combined with the EIT-based sensor developed in this thesis to give a sense of touch to the actuator in the form of EIT-based pressure mapping. Such a DEA-EIT actuator-sensor device has not been disseminated, so a patent towards this technology has been filed provisionally. This work has investigated the complications with integrating the two technologies such as the inherent trade off between pressure mapping resolution and actuation performance with the current construction of the device.

During the validation and characterisation of the DEA-EIT device this work shows the formation of an unintentional dielectric elastomer generator (DEG) and the ability to map and localise compressive events triggering the generation of energy. The energy generated by such a device in practice is dependent on the topology of the device. Using FEM studies, energy generation within the DEA-EIT device was estimated for a range of loading cases and compared to a simplified analytical approach.

Ultimately this work strived to create devices which mimic characteristics of human tissue, generally skin and skeletal muscle tissue. Through biomimicry we can start to develop devices that utilise the naturally evolved and optimised biological structures and functions which help us thrive on Earth. A non-exhaustive list of the characteristics of human muscle and skin tissue are compared to the characteristics of the soft robotic devices developed in this work in Table \ref{tab:biomimic-compare}.

\begin{table}
	\label{tab:biomimic-compare}
	\caption{A comparison of the biomimic qualities of note displayed in this work's devices.}
\begin{tabular}{|p{0.3\linewidth}|p{0.3\linewidth}|p{0.3\linewidth}|} 
	\textbf{Characteristic]} &
	\textbf{Biological Range} &
	\textbf{This work} \\ \hline
	Softness/Elasticity & 5 - 25 A / 10.4 - 118.4 kPa & ~10 A \ 98.1 - 132.5 kPa \\ \hline
	Spatial Resolution & 2 - 40 mm & 0.67 - 12 mm \\ \hline
	Temporal Resolution & 0.3 - 800 Hz & 0.1 Hz** \\ \hline
	Force Resolution & 0.6 - 122.6 mN & $\pm$ 78 - 81 mN \\ \hline
	Actuation Energy Density & 0.4 - 40 $\mathrm{J\cdot kg^{-1}}$ & 7.81 $\mathrm{J\cdot kg^{-1}}$ \\ \hline
	Actuation Strain & 0 - 30\% &	0 - 2.5\% \\ \hline
\end{tabular}
\end{table}

The softness, elasticity, spatial resolution, force resolution, actuation energy density values were within a comparable range for both this work's devices and the biological human equivalent. The spatial resolution in humans depends on the part of the body the test is subjected to due to the mechanoreceptors and other differences in skin tissue. Similarly the sensor spatial resolution is dependent on the location tested due primarily to current density distribution and material homogeneity. The actuation energy density estimate was done using a DEA-EIT sample at 10 kV using the thinnest DEA-EIT electrode configuration, see Appendix \ref{apdx:DEA_density}. Temporal resolution here has been given as 0.1 Hz however this value is dependent on the force resolution required and can change orders of magnitude depending on the desired force resolution. The actuation strain range of this work is lower than that of human skeletal muscle, however with refinement of the materials used as seen in literature it is hypothesised a DEA-EIT device could be created which matches and supersedes the actuation strain of human skeletal muscle.

% Insert references in table and above paragraph!

\section{Future Work}
To move the technology developed from Chapters \ref{chapter3} to \ref{chapter7} of this thesis into real-world applications and further enhance their performance, more research and development is required.

Although several methods of macro-modelling have been developed to represent a piezoresistive relationship in a conductive particle elastomer composite. There has been no evidence in literature of in-situ particle scale imaging of a CPEC specimen undergoing a known strain event. This in-situ imaging experiment could be done using imaging devices such as transmission electron microscopy (TEM) and small angle x-ray scattering (SAXS). This would show researchers how the particles move within a strained material. If using SAXS for nanoparticle imaging, an electrical current could also be applied to the material to determine whether an applied electric field changes the motion of the particles significantly and explains other electromechanical phenomena measured in the material. In parallel with this particle modelling, additional macro-level non-linear modelling needs to be pursued to mitigate time dependent phenomena slowing down the response time of the sensor.

A core issue for the CBSR composite material is quantifying the degree of material homogeneity in terms of carbon black particle and void dispersion and how to determine the piezoresistivity to calibrate based on non-homogeneity within the CBSR material volume. Depending on the location of a compressive load the detected change in resistance would vary massively. Determining this dispersion could be done with the same imaging methods mentioned above, TEM and SAXS.

Development of soft EIT-based pressure sensors have been limited by the time response of the material and measurement system in this work and previous works. Future work will investigate adding capacitive functionality to EIT-based pressure mapping by using a dielectric elastomer (DE) to aid capacitively shunt current so that not just the resistance of the device changes but also the capacitance. The hypothesis is that the non-linear time dependent effects of resistive sensing will be mitigated using this capacitive shunting. The viscoelasticity within the deformed DE is less prominent than that seen in the CBSR material giving the improved response time. As described in previous works \cite{Ellingham2024} there have been various attempts at created EIT-based pressure sensing devices. However, all of the state-of-the-art EIT-based pressure sensors rely on the change of resistance for pressure mapping. Zhang et. al \cite{Zhang2017} utilised capacitive shunting for EIT touch mapping on a variety of surfaces with conductive coatings. Their work used a human touch input to shunt electrical current through the person's contact point, essentially mapping localised changes in capacitance, but not applied surface pressure. Work from Reynolds-Smith \cite{Reynoldssmith1995,Reynoldssmith1999} showed the similarities and methods for Electric Field Tomography (EFT) as a method of localising touch and proximity of a grounded body. 

Another ubiquitous problem for an EIT-based pressure sensor is miniaturising the EIT reconstruction processor. The smallest processors used in literature are SoCs or FPGA implementations, however these are much slower than PCs in their current state \cite{ZamoraArellano2020,Kim2017,Takhti2019,Liu2019}. With the the current developments miniaturisation of parallel computing ICs, there are more possibilities to create an EIT-based pressure mapping sensor PCB with a processor, and/or FPGA or ASIC that can do complete both EIT data acquisition and reconstruction tasks.

The DEA-EIT device showed potential to also act as a DEG with with the simultaneous pressure localisation function. Simulations showed a range of energy generation quantities with a range of applied loads to determine the energy generation potential of real-world loading scenarios. The next step is to modify the existing electrical hardware to run the DEG sequence and measure the losses within the system to validate the potential for running EIT with the DEG generated power. Future work is going towards creating a DEA daughter board to fit the ERT circuit developed in this work, for further development of the DEA-EIT based solution.

\section{Concluding Statements}
Current commercially viable soft pressure mapping and soft actuator options are extremely limited so this thesis has explored soft sensor and actuator technology while maintaining specifications derived from biological human skin and muscle throughout the research and design process. In biology, we see the fusion of sensors in the form of receptors and actuators in the muscle cells within biological tissue. This work has taken inspiration from this sensor-actuator fusion and combined a soft pressure mapping sensor with a dielectric elastomer actuator to form a soft tissue-like multi-functional device.

Soft robotics is an emerging sub-market of the growing robotics and automation market. Currently we see soft sensor and actuator companies emerging, but adoption of the technology is slow. In production lines often there is a combination of human and robotic handling, especially in agricultural industries. This can lead to a dangerous and inefficient environment for human handlers to work in. In future applications, soft robotic solutions will be investigated to replace the human handlers in the production line for instances where safety and efficiency is desired. Another key field of application is the biomedical industry. It has been estimated that globally approximately 13 million amputations happen per year \cite{Yuan2023a}, and 0.9 million people per year have a spinal injury \cite{Ding2022}. Current prosthetic and assistive technology do not retain the characteristics of soft human skin and muscle tissue. Commercially viable soft sensor and actuator options are extremely limited, potentially indicating that the sensor and actuator technology requires further research to become more practical and reliable. Consequently, specific real-world applications were not pursued in this thesis; instead, the technology was characterised. The characterisation of device performance, and limitations help to affirm for future research that the technology is suitable for applications. Ultimately, the hope is that the biomimetic technology developed can be further researched and developed to improve human lives.

%Dream experiments include:
%- In-situ CBSR strain TEM imaging [x]
%- RNNs or other non-linear modelling techniques for modelling time dependent phenomena []
%- EIT with capacitive shunting AND/OR EFT-ERT MUX device [x]
%- DEG experiments to validate simulation [x]
%- develop an ASIC dedicated to gathering EIT data and completing inverse problems and pressure mapping co-processing?? []

% \lipsum

